Analysis and elimination of bottlenecks\\ in parallel algorithm for solving\\ global optimization problems

### 2022-04-15 2022/RSD/paper.tex
before submission

\usepackage{hyperref}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{amsmath} 
\usepackage{multirow} 
%
\begin{document}
%
\title{Analysis and elimination of bottlenecks in parallel algorithm for solving global optimization problems}
%
\titlerunning{Analysis and elimination of bottlenecks in parallel algorithm}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Ilya Lebedev \orcidID{0000-0002-8736-0652} \and
Denis Karchkov \orcidID{0000-0002-4164-5927}
}
%
\authorrunning{K. Barkalov, I. Lebedev, D. Karchkov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
\email{\{konstantin.barkalov,ilya.lebedev,karchkov\}@itmm.unn.ru}
}

@@

Constructing a configuration from a point $y$ is performed by a mapping of $y$ into a point $\bar{y} = F(y)$, where $F$ is an operator of constructing a configuration. It is constructed so that the direction $(\bar{y} - y)$ is that of the decreasing of the function $\varphi$ in the nearness of $y$. To describe the operator $F$, let us introduce the following notations: $e^i$ is the $i^{th}$ unit coordinate vector and $h$ is the parameter determining the magnitude of the coordinate translation. Then, the transition from $y$ to $\bar{y}$ (i.e. the algorithm of constricting the configuration $\bar{y} = F(y)$) is performed according to the following rules.

@@

Step 0. Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha = 2$ of the increasing of the step, e.g. $\alpha = 2$.

@@

\>\>\>\>\>\>\>\> if $h \leq \varepsilon$ then

\>\>\>\>\>\>\>\>\>\>\>\>  stop the search

@@

Step 4. Perform the step by direction $y^{k + 1} = t^k + \alpha (t^k - t^{k - 1})$ and go to Step 2.

The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. 
Step 4 was introduced in order to enable the method to increase the shift magnitude within a single iteration rapidly in the case when the point $t^k$ is located far away from the solution. At that, prior to make $y^{k+1}$ current point of the iteration, a configuration is constructed from the point $y^{k+1}$. By these means, in the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step by sample, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.

Finally, upon obtaining a value $\varphi(t^{k+1}) \geq \varphi(t^k)$ at Step 3, an attempt to run the algorithm again from the point $t^0=t^k$ (better than the one found) is undertaken. At that, if such an attempt hasn’t been undertaken yet, the parameter $h$ is not changed, but if it has been undertaken already, $h$ is cut in half first.

@@

Fig.~\ref{fig:RosenbrocNT} presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 2 to 30. Fig.~\ref{fig:RosenbrocSpeedup} presents corresponding speedups in iterations. The number of parallel trials $p$ varied from 1 to 16.

>>>>

\usepackage{hyperref}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{amsmath} 
\usepackage{multirow} 
%
\begin{document}
%
\title{Analysis and elimination of bottlenecks\\ in parallel algorithm for solving\\ global optimization problems}
%
\titlerunning{Analysis and elimination of bottlenecks in parallel algorithm}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Ilya Lebedev \orcidID{0000-0002-8736-0652} \and
Denis Karchkov \orcidID{0000-0002-4164-5927}
}
%
\authorrunning{K. Barkalov, I. Lebedev, D. Karchkov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
\email{\{konstantin.barkalov,ilya.lebedev,karchkov\}@itmm.unn.ru}
}

@@

Constructing a configuration from a point $y$ is performed by a mapping of $y$ into a point $\bar{y} = F(y)$, where $F$ is an operator of constructing a configuration. It is constructed so that the direction $(\bar{y} - y)$ is that of the decreasing of the function $\varphi$ in the nearness of $y$. To describe the operator $F$, let us introduce the following notations: $e^i$ is the $i^{th}$ unit coordinate vector and $h$ is the parameter determining the magnitude of the coordinate movement. Then, the transition from $y$ to $\bar{y}$ (i.e. the algorithm of constricting the configuration $\bar{y} = F(y)$) is performed according to the following rules.

@@

Step 0. Set the starting point $t^0$, the parameter of the method $\varepsilon > 0$, the parameter of constructing the configuration $h\gg\varepsilon$, and the parameter $\alpha$ of the increasing of the step, e.g. $\alpha = 2$.

@@

\>\>\>\>\>\>\>\> if $h \leq \varepsilon$ then stop the search

@@

Step 4. Move by direction $y^{k + 1} = t^k + \alpha (t^k - t^{k - 1})$ and go to Step 2.

The meaning of the actions performed within the steps 2, 3, and 4 can be explained as follows. 
Step 4 was introduced in order to enable the method to rapidly increase the shift magnitude within a single iteration in the case when the point $t^k$ is located far away from the solution. At that, prior to make $y^{k+1}$ current point of the iteration, a configuration is constructed from the point $y^{k+1}$. In the case of obtaining a point with $\varphi(t^{k+1})< \varphi(t^k)$, the next step, in general, will be performed in a changed direction (relative to previous one) that allows adapting the method to variations in the function relief.

Finally, upon obtaining a value $\varphi(t^{k+1}) \geq \varphi(t^k)$ at Step 3, an attempt to run the algorithm again from the point $t^0=t^k$ (better than the one found) is undertaken. At that, if such an attempt hasn't been undertaken yet, the parameter $h$ is not changed, but if it has been undertaken already, $h$ is cut in half first.

@@

Fig.~\ref{fig:RosenbrocNT} presents the averaged numbers of iterations, which Hooke-Jeeves and BFGS methods executed for minimization of Rosenbrock function with the dimensionalities from 5 to 30. Fig.~\ref{fig:RosenbrocSpeedup} presents corresponding speedups in iterations. The number of parallel trials $p$ varied from 1 to 16.
